---
title: "Assignment 4"
output: html_notebook
author: "Maya Stein"
date: "`r Sys.Date()`"
---

# Set Up

Load the libraries
```{r}
library(tidyverse)
library(dplyr)
library(psymetadata)
```
Look at the data set
```{r}
?barroso2021
```
```{r}
barroso2021
```

First I will explore the number of students with low vs high math ability per each grade
```{r}
barroso2021 |> 
  group_by(grade) |> 
  count(low_ability)
```

It seems there are no students identified as having low math ability in grades 1, 3, 4, and 6. Now I will confirm if this is accurate

Assess grades 1, 3, 4, and 5
```{r}
barroso2021 |> 
  filter(grade==1 | grade==3 | grade==4 | grade==6) |> 
  distinct(low_ability)
```

Assess grades 2 and 5
```{r}
barroso2021 |> 
  filter(grade==2 | grade==5) |> 
  distinct(low_ability)
```

I now need to observe the sample size for each grade
```{r}
barroso2021 |> 
  group_by(grade) |> 
  summarize(n=n())
```
Grade 5 has the biggest sample, as well as students for both high and low math ability conditions. Grade 6 has the lowest sample size, which likely explains the lack of samples for students with low math abilities. The absence/presence of low math ability students in the other grades is not well understood at this time. This will likely play a role in observations of effect size.

